---
title: "Reproducing the fitted inverse covariance matrix used in the  paper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{smoking-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(sparse.inv.cov)

```



# pre-processing the smoking data

The data used to illustrate the method implemented in mvama was discussed in: Spira, A. et
al. (2004). Effects of cigarette smoke on the human airway epithelial cell transcriptome. Proc. Natl
Acad. Sci. USA, 101, 10143-10148.  The raw data may be obtained from the NCBI at
\url{http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE994}. Only CEL files corresponding to
'smokers' and 'never smokers were used in the paper; these correspond to Accession IDs GSM15684-
15717 and GSM15718-15740, respectively.  Once the compressed CEL files (GSM15XYZ.CEL.gz) have been
downloaded to a directory, they can be read in and rma-corrected as follows: usiing the following
Bioconductor (http://www.bioconductor.org) packages:

```{r, echo=TRUE,eval=FALSE}
require(affy)
require(limma)
require(hgu133acdf)
data <- ReadAffy(celfile.path = "/path_to_CEL_files", compress = TRUE)

```
Carry out rma correction
```{r, echo=TRUE,eval=FALSE}
rma.data <- rma(data)
```
Note that this is equivalent to using the function 'expresso' as follows:
```{r, echo=TRUE,eval=FALSE}
rma.data <- expresso(data,
bgcorrect.method = "rma",
normalize.method = "quantiles",
pmcorrect.method = "pmonly",
summary.method = "medianpolish")
```

And finally,
```{r, echo=TRUE,eval=FALSE}
rma.data <- t(exprs(rma.data))
```
which gives a 57 x 22283 matrix where the first 34 rows correspond
to 'smokers', and the next 23 correspond to 'never-smoked'. Consequently, the vector of class labels y can be
generated as
```{r, echo=TRUE,eval=FALSE}
y <- rep(1:2, times = c(34, 23))

```


# create the R workspace smoking.RData
The pre-processed data set is now included in this package
```{r, echo=TRUE,eval=FALSE}
library(sparse.inv.cov)
data(smoking,package="sparse.inv.cov")
names(smoking)
#[1] "y"   "rma"
head(smoking$y)
# "smokers" "smokers" "smokers" "smokers" "smokers" "smokers"
y <- rep(1:2, times = c(34, 23))
smoking$rma[1:3,1:5]
#                1007_s_at  1053_at   117_at   121_at 1255_g_at
#GSM15684.CEL.gz  9.553334 4.182281 5.552527 7.984070  4.267715
#GSM15685.CEL.gz 10.025818 4.401058 5.120846 7.879789  3.806086
#GSM15686.CEL.gz  9.640811 4.327433 5.682144 7.957257  3.984375

X<-smoking$rma

```

# Compute the residual matrix after removing mean structure.

generate the design matrix
```{r, echo=TRUE,eval=FALSE}
D<-matrix(0,nrow=57,ncol=2)
D[,1]<-as.numeric(y==2)
D[,2]<-as.numeric(y==1)

R<-lm(X~D-1)$residuals

```
# Determine zero pattern

```{r, echo=TRUE,eval=FALSE}
#library(Matrix)
library(sparse.inv.cov)

```
single run if you have the time and patience
```{r, echo=TRUE,eval=FALSE}
res<-get.neighbs(R,kmax=3)
a01 <- res$a01

```
use modified BIC with g=1.
The computation is easily paralllelizable if we have mutliptle processons. 
```{r, echo=TRUE,eval=FALSE}
vignette(parallel_processing)

```
gives further details.

<!-- ## split it over more processors,  -->
<!-- in this case 5,  Source the scripts nb.script1.r to nb.script5.r on separate processors. -->
<!-- These scripts may need to be edited so that the load and save statements -->
<!-- read and write from a specified directory. After these jobs have finished, running R -->
<!-- in the common directory and sourcing the script nb.collate.r produces the list -->
<!-- bres10. From this set -->
<!-- ```{r, echo=TRUE,eval=FALSE} -->
<!-- a<-bres10$a -->
<!-- ``` -->

# Fit the inverse covariance matrix

```{r, echo=TRUE,eval=FALSE}
    res<-hd.covsel(R,a,nsamp=57,corr=TRUE,eps=.01,m=200)

```
depending on your processor speed the fit could take between 1 and 2 days
larger values of m might reduce this

# compute inverse covariance matrix from fitted inverse correlation matrix
assumes residual matrix R is available as well as Matrix library
```{r, echo=TRUE,eval=FALSE}
sd<-apply(R,2,var)56/
sd<-sd^0.
d<-Diagonal(length(sd),1/sd)
sinv<-d%%res$si%*%d

```
