---
title:  "Typical workflow for using the sparse.inv.cov R library"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-stuff}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(sparse.inv.cov)
```
It is recommended that the reader runs the code here in R and also consults the help
files in the library.

# Get data
For illustrative purposes here we simulate some data with known structure. First
we generate the $\Sigma^{-1}$ matrix. 

```{r, echo=TRUE,eval=FALSE}
sinv<-diag(5)/
sinv[1,2]=sinv[2,3]=sinv[3,4]=sinv[4,5]=0.
sinv<-sinv+t(sinv)
sinv<-as(sinv,"dsCMatrix")
```
Then we generate data with mean zero and inverse covariance matrix sinv
```{r, echo=TRUE,eval=FALSE}
res<-Simulate(sigma.inv=sinv,M=NULL,nsamp=1000,nsim=1,seed=1)
X<-res$data[[1]]
```
We mean correct each column
```{r, echo=TRUE,eval=FALSE}
X<-scale(X,center=TRUE,sc=FALSE)
```
more generally regress each column of X on the design matrix and replace it with
the residuals. We check the means
```{r, echo=TRUE,eval=FALSE}
colMeans(X)
```
check structure in observed sample inverse and compare with the true inverse
```{r, echo=TRUE,eval=FALSE}
solve(crossprod(X)/1000)
sinv
```
# Find zero pattern

there can be at most 4 neighbours for each variable here. We
generally want m/n <1/20 where m is the number of neighbours of a variable
```{r, echo=TRUE,eval=FALSE}
res<-get.neighbs(X,kmax=4)
```
For a discussion of the parameter kmax see Kiiveri(2011).

## example of splitting the job over two processors to speed things up
on processor 1
```{r, echo=TRUE,eval=FALSE}
res1<-get.neighbs(X,kmax=4,cols=1:2)
save(res1,file="mydir/res1.RData")
```
and on processor 2
```{r, echo=TRUE,eval=FALSE}
res2<-get.neighbs(X,kmax=4,cols=3:5)
save(res2,file="mydir/res2.RData")
```
then combine results
```{r, echo=TRUE,eval=FALSE}
load("mydir/res1.RData")
load("mydir/res2.RData")
```
get incidence matrix for all variables
```{r, echo=TRUE,eval=FALSE}
a00<-res1$a00+res2$a00
```
 similar results apply for a05 a01 etc.
The nonzero entries in a00 could be reset to 1 but typically it doesn’t matter for
further processing. The generalisation to more than 2 processors should be clear

# Fit covariance selection model using default parameters
use res\$a00 here (the usual BIC) because n is much greater than p. When p>>n use
res$a01 from the previous step.
```{r, echo=TRUE,eval=FALSE}
tmp<-hd.covsel(X, res$a00, nsamp=nrow(X),corr=FALSE)
```
look at fitted inverse matrix
```{r, echo=TRUE,eval=FALSE}
tmp$sigma.inv
```
look at fitted covariance matrix in positions where sigma.inv is not zero
```{r, echo=TRUE,eval=FALSE}
tmp$sigma
```

Reference
Kiiieri(2011) Multivariate analysis of microarray data: differential expression and
differential connection. BMC Bioinformatics

